Physcraper: A Python package for continually updated phylogenetic trees using the Open Tree of Life

### 2021-05-01 sources/bmc_template_submission/bmc_article.tex
fixing 'unified' typos

Data such as
geographical location, fossil ranges, and genetic and phenotypic information increasingly
available in public biological databases constitute an amazing
resource for scientific discovery \cite{baxevanis2015importance}.
One of the main challenges for automatic
integration of data across biological databases are varying taxonomic idiosyncrasies.
To address this challenge, the Open Tree of Life project (OpenTree)
created a unified taxonomy for automatic taxonomic name standardization, by integrating taxonomic
data from several resources \cite{rees2017automated}, including the USA National
Center for Biodiversity Information (NCBI) taxonomy \cite{federhen_ncbi_2012, schoch_ncbi_2020},
and the Global Biodiversity Information Facility (GBIF) \cite{gbif_secretariat_gbif_2019}, among many others.
OpenTree's unfified taxonomy, along with tools and methods to manage it, are available
as Application Programming Interfaces (APIs),
which are implemented as open access services for the general public \cite{opentreeAPIs}.
Physcraper leverages on existing OpenTree's unified taxonomy APIs to automatically
standardize taxon names in any phylogeny, providing
a framework for connection of updated phylogenies with data from any biological database.

@@

Taxon names in the input phylogeny must be standardized to OpenTree's unfified taxonomy
\cite{ott3.2} using OpenTree's bulk Taxonomic Name Resolution Service
(TNRS) tool \cite{TNRStool}. Users can upload their
own phylogeny, or choose from among the 2, 950 curated phylogenies stored in OpenTree's
Phylesystem database \cite{phylesystemGithub} that also have
alignments available on the TreeBASE database \cite{treebase_website, supertreebase}.

@@

BLAST is performed using the blastn algorithm \cite{camacho2009blast}
implemented in BioPython 1.71 \cite{cock2009biopython} NCBIWWW module \cite{ncbiwww}
modified to accept an alternative BLAST address.
Each sequence in the alignment is BLASTed once against the GenBank database.
Matching sequences are filtered and excluded from the analysis if they 1) are not in the search taxon;
2) have an e-value above the cutoff (default to 0.00001); 3) fall outside a minimum
and maximum sequence length threshold, defined as a proportion of the average sequence length without
gaps of all sequences in the input alignment (default values of 80\% and 120\%,
respectively); 4) or, if they are either identical to or shorter than an existing
sequence in the input alignment and they represent the same taxon in NCBI's or
OpenTree's unififed taxonomy.
By default, an arbitrary maximum number of 5 sequences per taxon are chosen at random
from the set of matching sequences that passed the filtering step.

>>>>

Data such as
geographical location, fossil ranges, and genetic and phenotypic information increasingly
available in public biological databases constitute an amazing
resource for scientific discovery \cite{baxevanis2015importance}.
One of the main challenges for automatic
integration of data across biological databases are varying taxonomic idiosyncrasies.
To address this challenge, the Open Tree of Life project (OpenTree)
created a unified taxonomy for automatic taxonomic name standardization, by integrating taxonomic
data from several resources \cite{rees2017automated}, including the USA National
Center for Biodiversity Information (NCBI) taxonomy \cite{federhen_ncbi_2012, schoch_ncbi_2020},
and the Global Biodiversity Information Facility (GBIF) \cite{gbif_secretariat_gbif_2019}, among many others.
OpenTree's unified taxonomy, along with tools and methods to manage it, are available
as Application Programming Interfaces (APIs),
which are implemented as open access services for the general public \cite{opentreeAPIs}.
Physcraper leverages on existing OpenTree's unified taxonomy APIs to automatically
standardize taxon names in any phylogeny, providing
a framework for connection of updated phylogenies with data from any biological database.

@@

Taxon names in the input phylogeny must be standardized to OpenTree's unified taxonomy
\cite{ott3.2} using OpenTree's bulk Taxonomic Name Resolution Service
(TNRS) tool \cite{TNRStool}. Users can upload their
own phylogeny, or choose from among the 2, 950 curated phylogenies stored in OpenTree's
Phylesystem database \cite{phylesystemGithub} that also have
alignments available on the TreeBASE database \cite{treebase_website, supertreebase}.

@@

BLAST is performed using the blastn algorithm \cite{camacho2009blast}
implemented in BioPython 1.71 \cite{cock2009biopython} NCBIWWW module \cite{ncbiwww}
modified to accept an alternative BLAST address.
Each sequence in the alignment is BLASTed once against the GenBank database.
Matching sequences are filtered and excluded from the analysis if they 1) are not in the search taxon;
2) have an e-value above the cutoff (default to 0.00001); 3) fall outside a minimum
and maximum sequence length threshold, defined as a proportion of the average sequence length without
gaps of all sequences in the input alignment (default values of 80\% and 120\%,
respectively); 4) or, if they are either identical to or shorter than an existing
sequence in the input alignment and they represent the same taxon in NCBI's or
OpenTree's unified taxonomy.
By default, an arbitrary maximum number of 5 sequences per taxon are chosen at random
from the set of matching sequences that passed the filtering step.
